### 2024-01-02 chapters/conclusion.tex
Polish conclusion

\textbf{\Cref{chp:introduction}} introduced the motivational cases of automated medical coding and stroke recognition and used them to exemplify the importance of out-of-distribution detection, and, by extension, representation learning. In the context of these cases, we discussed possible machine learning system designs for decision support and considered potential sources of uncertainty and ideal model behavior. 
The cases form a reference point for the thesis as a whole, connecting its contributions within out-of-distribution detection and representation learning back to practical applications. 
In this conclusion, we will review the studies presented in previous chapters in the context of the discussion of \cref{chp:discussion}, the recent progress in the field, and point to interesting directions of future research.
As \textbf{\cref{chp:main-contributions}} already details the contributions made by this thesis, we will not reiterate them in detail in this conclusion.

@@

\vspace{1em}
\textbf{\Cref{chp:technical-background}} provided technical background that could only be covered briefly by the individual studies. 
We first introduced uncertainty as a concept in the context of information and probability theory (\cref{sec:uncertainty-information-theory}). We then defined the task of out-of-distribution detection and reviewed existing work on the problem (\cref{sec:out-of-distribution-detection}). Finally, we provided technical background for variational autoencoders (\cref{sec:variational-autoencoders}).

\vspace{1em}
\textbf{\Cref{chp:paper-hierarchical}} showed how hierarchical variational autoencoders can fail at likelihood"-based OOD detection due to an overemphasis on low-level features that generalize between different data distributions. 
In other words, the distributions of latent representations of high-dimensional data overlap, especially for latent variables low in the hierarchy.
By exploiting that VAEs tend to learn more abstract features at latent variables high in the hierarchy, we were able to define a likelihood-ratio score that focused more on features unlikely to be shared between datasets and performed much better for OOD detection. 
In line with our discussion in \cref{sec_discussion:representation-learning-with-vaes}, these findings show that VAEs can indeed learn useful latent representations, although good performance, at least for OOD detection, might require selecting an appropriate subset of latent representations to use. 
Similar variations among features learned at different layers have also been identified for self"=supervised models and speech representations \parencite{pasad_layerwise_2021}. 
To obtain good downstream task performance, this suggests that, besides seeking to improve representations overall, future research effort should also be directed towards finding good methods for selecting relevant subsets of latent variables in a given hierarchical VAE for a certain task.
% - VAEs can learn useful representations, but it can be hard to impose the appropriate constraints that enable this.
% - VAEs can learn representations that are useful for OOD detection, but good performance might require selecting an appropriate subset of feature dimensions to use for discriminating.
% Our work shows that a promising direction of future research is into how to select which dimensions are the most discriminatory.

@@

\vspace{1em}
\textbf{\Cref{chp:paper-modelagnostic}} took a different approach to OOD detection than \cref{chp:paper-hierarchical} and focused on developing a model-agnostic method. 
We showed that by phrasing OOD detection as a statistical testing problem and combining different tests, orthogonal properties of the individual tests could be leveraged to improve the OOD detection performance over any single test.

The formulation of OOD detection as a statistical test also allows for better guarantees for such systems in practice. For instance, as also discussed in \cref{chp:paper-modelagnostic}, the statistical framework enables false positive rate control which is a valuable property in many practical applications especially if they involve high-risk actions such as in medical decision support.

@@

\vspace{1em}
\textbf{\Cref{chp:paper-brief}} provided an overview of unsupervised neural speech representation learning. Such approaches have recently matched supervised methods on many tasks and represent a significant advance in low-resource settings, such as speech recognition for minority languages. 
We found that for the purpose of learning good representations in an unsupervised manner, self-supervised learning seems to have better inductive biases, or at least pose a more forgiving learning problem, than do VAEs. As discussed in \cref{chp:paper-brief,sec_discussion:representation-learning-with-vaes}, this likely relates more to inductive biases imposed by implicit constraints in the optimization problem and architecture than to the underlying formalism. 
For instance, in discussion about the weaknesses of VAE-based approaches (\cref{sec_discussion:representation-learning-with-vaes}) we concluded that their challenges could not be directly attributed to the maximum marginal likelihood objective. Indeed, the masked pre-training objective widely used for successful self"=supervised methods also corresponds to a maximum marginal likelihood objective \parencite{moreno-munoz_masked_2023}.

This also leaves potential for future work to improve the ability of VAEs to learn useful representations. As discussed in \cref{chp:paper-brief,sec_discussion:representation-learning-with-vaes}, promising approaches include adopting masked objectives for VAE training, improved architectural designs to impose better inductive biases, and incorporating advances in gradient estimators more widely in the literature \parencite{rainforth_tighter_2019,roeder_sticking_2017,tucker_doubly_2019,bauer_generalized_2021}, especially works on large VAE models \parencite{maaloe_biva_2019,vahdat_nvae_2020,child_very_2021}.

@@

\vspace{1em}
\textbf{\Cref{chp:paper-benchmarking}} conducted a comprehensive evaluation of stochastic and deterministic generative models, focusing on their model likelihood. 
The chapter also introduced a novel hierarchical VAE type model for speech, by drawing inspiration from the Clockwork VAE \parencite{saxena_clockwork_2021}. 
Despite the supposed limitations of VAEs for representation learning as compared to self"=supervised methods (\cref{chp:discussion}), hierarchical VAE models for sequence data that operate on multiple temporal scales, provide a natural framework for encoding distinct feature categories. 
For instance, pronunciation features may be learned at lower layers, speaker identity at upper layers, and semantic features in intermediate layers. Despite successful attempts at isolating speaker identity from content in some existing work \parencite{hsu_unsupervised_2017}, developing VAE models with the capacity to learn a deep hierarchy of features for speech persists. The model presented in \cref{chp:paper-benchmarking} is an attempt at this challenge.
% Despite the current limitations, VAEs possess a unique ability to learn a distribution over the training data, requiring them to encode diverse aspects of the dataset. This inclusivity, while potentially redundant for certain tasks, proves advantageous for others that benefit from a wide array of features.

\vspace{1em}
\textbf{\Cref{chp:paper-automated}} examined existing works on automated medical coding and found that, in any cases, training was suboptimal, and evaluation standards were biased. We performed a revised comparison of the selected models and provided updated conclusions on the relative performance of models, and the impact of rare codes and long discharge note documents.

The practical impact of the work of \cref{chp:paper-automated}, and the work in the field of medical coding in general, depend on a number of factors. 
Much of the field has focused on the MIMIC datasets which originate from the emergency department and ICU of a single hospital. These datasets have enabled much of the progress in the field, but their singular data source also reduces the generality of the derived results and risks biasing the directions of research deemed most impactful \parencite{tengReviewDeepNeural2022, venkateshAutomatingOverburdenedClinical2023,johnsonMIMICIIIFreelyAccessible2016,johnsonMIMICIVFreelyAccessible2023}. 
Furthermore, the complexity and multi-label nature of medical coding has lead to a high prevalence of label errors in MIMIC-III \parencite{searleExperimentalEvaluationDevelopment2020}. While difficult to remove entirely, having multiple diverse datasets would also alleviate the risk of biases in such errors leading to misguided conclusions. Besides gathering more data, as also mentioned in \cref{chp:paper-automated}, evaluation methods, such as human-in-the-loop, could be useful to increase the reliability of results.

A weakness of medical coding models is that performance varies widely between classes. The long-tailed distribution of code-frequency is particularly challenging and leads to underperformance on rare codes. Practical applications could still benefit from such models though. By limiting model predictions to a subset of codes for which they perform well, practitioners could focus on harder to code cases. This selection of cases in practical applications of medical coding could also benefit from research into selective prediction \parencite{geifman_selective_2017}. 
Although pre"=training is an obvious approach to improve performance when little data is available, the improvements observed by using pre"=trained models for medical coding have been limited \parencite{jiDoesMagicBERT2021,gaoLimitationsTransformersClinical2021,michalopoulosICDBigBirdContextualEmbedding2022,pascualBERTbasedAutomaticICD2021,zhangBERTXMLLargeScale2020} compared to the effect of pre"=training in other domains \parencite{mohamed_selfsupervised_2022, linPretrainedTransformersText2021,baevski_wav2vec_2020,devlin_bert_2018,dosovitskiy_image_2021}. This suggests an untapped potential for future research into more targeted pre"=training for medical coding.

While our work in \cref{chp:paper-automated} focused on unimodal medical coding models, it is highly likely that future state-of-the-art models will augment discharge summaries with multi-modal inputs such as medical code descriptions, synonyms, and hierarchies \parencite{kimReadAttendCode2021, mullenbachExplainablePredictionMedical2018, vuLabelAttentionModel2020, caoHyperCoreHyperbolicCograph2020, baoMedicalCodePrediction2021, yuanCodeSynonymsMatter2022,caoHyperCoreHyperbolicCograph2020, xieEHRCodingMultiscale2019}. 
A main reason for this is that coding standards are updated regularly. The ICD standard for instance sees revisions and new codes added yearly, with local adoption following national guidelines \parencite{centerfordiseasecontrolandpreventioncdc_international_2023}. 
Leveraging such modalities will enable adapting models to updated coding standards without having to gather large amounts of data.

@@

In \cref{sec: discussion-stroke-recognition-uncertainty}, we discussed how calibrating the predictive uncertainty of the stroke model is likely to be necessary to enable sustained use of such a model in practice. In \cref{fig_discussion:retrospective-paper-f1-performance-vs-predicted-probability}, we noted that, as expected, model performance measured by F1-score increased with increasing model certainty which underlines the likely usefulness of uncertainty estimates in matching practitioner expectations of the predictive performance. 
Nonetheless, basic metrics of model performance might still be obstacles for its practical usefulness. Specifically, the rarity of stroke cases lead to a false positive rate and precision that are likely to induce alarm fatigue among its users. Similar effects are likely to have influenced the practical impact of a similar system for cardiac arrest detection which showed significant improvements retrospectively \parencite{cite14} which, although matched by the model in a prospective study, did not ultimately result in improved call-taker performance \parencite{cite15}.

>>>>

\textbf{\Cref{chp:introduction}} introduced the motivational cases of automated medical coding and stroke recognition and used them to exemplify the importance of out-of-distribution detection, and, by extension, representation learning. In the context of these cases, we discussed possible machine learning system designs for decision support and considered potential sources of uncertainty and ideal model behavior. 
The cases form a reference point for the thesis as a whole, connecting its contributions within out-of-distribution detection and representation learning back to practical applications. 
In this conclusion, we will review the studies presented in previous chapters in the context of the discussion of \cref{chp:discussion} and the recent progress in the field, and point to interesting directions of future research. 
As \textbf{\cref{chp:main-contributions}} already details the contributions made by this thesis, we will not reiterate them in detail in this conclusion.

@@

\vspace{1em}
\textbf{\Cref{chp:technical-background}} provided in-depth technical background that could only be covered briefly by the individual studies. 
We first introduced uncertainty as a concept in the context of information and probability theory (\cref{sec:uncertainty-information-theory}). We then defined the task of out-of-distribution detection and reviewed existing work on the problem (\cref{sec:out-of-distribution-detection}). Finally, we provided technical background for variational autoencoders (\cref{sec:variational-autoencoders}).

\vspace{1em}
\textbf{\Cref{chp:paper-hierarchical}} showed how hierarchical variational autoencoders can fail at likelihood"-based OOD detection due to an overemphasis on low-level features that generalize between different data distributions. 
In other words, the latent representations of high-dimensional data from different distributions overlap, especially for latent variables low in the hierarchy.
By exploiting that VAEs tend to learn more abstract features at latent variables high in the hierarchy, we were able to define a likelihood-ratio score that focused more on features unlikely to be shared between datasets and performed much better for OOD detection. 
In line with our discussion in \cref{sec_discussion:representation-learning-with-vaes}, these findings show that VAEs can indeed learn useful latent representations, although good performance, at least for OOD detection, might require selecting an appropriate subset of latent representations to use. 
Similar variations among features learned at different layers have also been identified for self"=supervised models and speech representations \parencite{pasad_layerwise_2021}. 
To obtain good downstream task performance, this suggests that, besides seeking to improve representations overall, future research effort should also be directed towards finding good methods for selecting relevant subsets of latent variables in a given hierarchical VAE for a certain task.
% - VAEs can learn useful representations, but it can be hard to impose the appropriate constraints that enable this.
% - VAEs can learn representations that are useful for OOD detection, but good performance might require selecting an appropriate subset of feature dimensions to use for discriminating.
% Our work shows that a promising direction of future research is into how to select which dimensions are the most discriminatory.

@@

\vspace{1em}
\textbf{\Cref{chp:paper-modelagnostic}} took a different approach to OOD detection than \cref{chp:paper-hierarchical} and focused on developing a model-agnostic method. 
We showed that by phrasing OOD detection as a statistical testing problem and combining different tests, orthogonal properties of the individual tests could be leveraged to improve the OOD detection performance over any single test.
%
The formulation of OOD detection as a statistical test also allows for better guarantees for such systems in practice. For instance, as also discussed in \cref{chp:paper-modelagnostic}, the statistical framework enables false positive rate control which is a valuable property in many practical applications especially if they involve high-risk actions such as in medical decision support.

@@

\vspace{1em}
\textbf{\Cref{chp:paper-brief}} provided an overview of unsupervised neural speech representation learning. Such approaches have recently matched supervised methods on many tasks and represent a significant advance in low-resource settings, such as speech recognition for minority languages. 
We found that for the purpose of learning good representations in an unsupervised manner, self-supervised learning seems to have better inductive biases, or at least pose a more forgiving learning problem, than do VAEs. As discussed in \cref{chp:paper-brief,sec_discussion:representation-learning-with-vaes}, this likely relates more to inductive biases imposed by implicit constraints in the optimization problem and architecture than to the underlying formalism. 
For instance, in discussion about the weaknesses of VAE-based approaches (\cref{sec_discussion:representation-learning-with-vaes}) we concluded that their challenges could not be directly attributed to the maximum marginal likelihood objective. Indeed, the masked pre-training objective widely used for successful self"=supervised methods has also been identified to correspond to a maximum marginal likelihood objective \parencite{moreno-munoz_masked_2023}.

This also leaves potential for future work to improve the ability of VAEs to learn useful representations. As discussed in \cref{chp:paper-brief,sec_discussion:representation-learning-with-vaes}, promising approaches include adopting masked objectives for VAE training, improving architectural designs to impose better inductive biases, and incorporating advances in gradient estimators more widely \parencite{rainforth_tighter_2019,roeder_sticking_2017,tucker_doubly_2019,bauer_generalized_2021}, especially works on large VAE models \parencite{maaloe_biva_2019,vahdat_nvae_2020,child_very_2021}.

@@

\vspace{1em}
\textbf{\Cref{chp:paper-benchmarking}} conducted a comprehensive evaluation of stochastic and deterministic generative models, focusing on their model likelihood. 
The chapter also introduced a novel hierarchical VAE type model for speech, by drawing inspiration from the Clockwork VAE \parencite{saxena_clockwork_2021}. 
Despite the limitations of VAEs for representation learning as compared to self"=supervised methods (\cref{chp:discussion}), for sequence data, hierarchical VAE models that operate on multiple temporal scales provide a natural framework for encoding distinct feature categories. 
For instance, pronunciation features might be learned at lower layers, speaker identity at upper layers, and semantic features in intermediate layers. Despite successful attempts at isolating speaker identity from content in some existing work \parencite{hsu_unsupervised_2017}, developing VAE models with the capacity to learn a deep hierarchy of features for speech persists. The model presented in \cref{chp:paper-benchmarking} is an attempt at this challenge.
% Despite the current limitations, VAEs possess a unique ability to learn a distribution over the training data, requiring them to encode diverse aspects of the dataset. This inclusivity, while potentially redundant for certain tasks, proves advantageous for others that benefit from a wide array of features.

\vspace{1em}
\textbf{\Cref{chp:paper-automated}} examined existing works on automated medical coding and found that in many cases training was suboptimal and evaluation standards were biased. We performed a revised comparison of the selected models and provided updated conclusions on the relative performance of models, and the impact of rare codes and long discharge note documents.

The practical impact of the work of \cref{chp:paper-automated}, and the work in the field of medical coding in general, depend on a number of factors. 
Much of the field has focused on the MIMIC datasets which originate from the emergency department and ICU of a single hospital. These datasets have enabled much of the progress in the field, but their singular data source also reduces the generality of the derived results and risks biasing the directions of research deemed most impactful \parencite{tengReviewDeepNeural2022, venkateshAutomatingOverburdenedClinical2023,johnsonMIMICIIIFreelyAccessible2016,johnsonMIMICIVFreelyAccessible2023}. 
Furthermore, the complexity and multi-label nature of medical coding has lead to a high prevalence of label errors in MIMIC-III \parencite{searleExperimentalEvaluationDevelopment2020}. While difficult to remove entirely, having multiple diverse datasets would also alleviate the risk that the biases of such errors lead to misguided conclusions. Besides gathering more data, improved evaluation methods, such as human-in-the-loop, could be useful to increase the reliability of results.

A common weakness of medical coding models is that performance varies widely between classes. The long-tailed distribution of code-frequency is particularly challenging and leads to underperformance on rare codes. Practical applications could still benefit from such models though. By limiting model predictions to a subset of codes for which they perform well, practitioners could focus on harder to code cases. This selection of cases in practical applications of medical coding could also benefit from research into selective prediction \parencite{geifman_selective_2017}. 
Although pre"=training is an obvious approach to improve performance when little data is available, compared to the effect of pre"=training in other domains \parencite{mohamed_selfsupervised_2022, linPretrainedTransformersText2021,baevski_wav2vec_2020,devlin_bert_2018,dosovitskiy_image_2021}, the improvements observed by using pre"=trained models for medical coding have been limited \parencite{jiDoesMagicBERT2021,gaoLimitationsTransformersClinical2021,michalopoulosICDBigBirdContextualEmbedding2022,pascualBERTbasedAutomaticICD2021,zhangBERTXMLLargeScale2020}. This suggests an untapped potential for future research into more targeted pre"=training for medical coding.

While our work in \cref{chp:paper-automated} focused on unimodal medical coding models, it is highly likely that future state-of-the-art models will augment discharge summaries with multi-modal inputs such as medical code descriptions, synonyms, and hierarchies \parencite{kimReadAttendCode2021, mullenbachExplainablePredictionMedical2018, vuLabelAttentionModel2020, caoHyperCoreHyperbolicCograph2020, baoMedicalCodePrediction2021, yuanCodeSynonymsMatter2022,caoHyperCoreHyperbolicCograph2020, xieEHRCodingMultiscale2019}. 
This is particularly useful since coding standards are updated regularly. 
The ICD standard for instance sees revisions and new codes added yearly, with local adoption following national guidelines \parencite{centerfordiseasecontrolandpreventioncdc_international_2023}. 
Leveraging such modalities will enable adapting models to updated coding standards without having to gather large amounts of new data.

@@

In \cref{sec: discussion-stroke-recognition-uncertainty}, we discussed how calibrating the predictive uncertainty of the stroke model is likely to be necessary to ensure sustained use of such a model in practice. In \cref{fig_discussion:retrospective-paper-f1-performance-vs-predicted-probability}, we noted that, as expected, model performance measured by F1-score increased with increasing model certainty which underlines the likely usefulness of uncertainty estimates in better matching practitioner expectations of the predictive performance. 
Nonetheless, basic metrics of model performance might still be obstacles for its practical usefulness. Specifically, the rarity of stroke cases lead to relatively high false positive rate and low precision, likely to induce alarm fatigue among its users. Similar effects are likely to have influenced the practical impact of a similar system for cardiac arrest detection which also showed significant improvements in a retrospective study \parencite{cite14}. Although the model later matched the retrospective results in a prospective study, it did not ultimately result in improved call-taker performance \parencite{cite15}.
